---
title: "How to quickly wrangle, clean, and summarize your CoralNet data"
author: "lkm"
date: "March 14, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# What you need to get started:  

* your exported metadata file  
* your exported annotations file (this is your data) 
* labelset .csv (provided in the github repo)  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) #install.packages(tidyverse)
library(janitor)

#check your working directory
getwd()

#Note: Change your working directory, if necessary. The working directly should be the folder your data is stored in.
#setwd()
```

#Constants  
This is where users can define variable names for imported datasets, especially if they are specific to a filepath. Here you will create variables called metadata and data, using your specific filepath location
```{r constants}

## Example code to add your own metadata and data variables, remove the # to run:

#metadata <- "add filename path"
#read_csv(metadata variable from above)

#data <- "add filename path"
#read_csv(data variable from above)

## Here we are reading in the metadata and data files from the impage_point_count_wrangle github repo:

metadata <- read_csv("https://raw.githubusercontent.com/Lmudge13/image_point_count_wrangling/master/sample_metadata_coralnet.csv")

data <- read_csv("https://raw.githubusercontent.com/Lmudge13/image_point_count_wrangling/master/sample_data_annotations_simple_coralnet.csv")


# ALL users can import this labelset using the follwing code:
labelset <- read_csv("https://raw.githubusercontent.com/Lmudge13/image_point_count_wrangling/master/coralnet_labelset.csv")

```

#Functions
```{r functions}

## add code for functions here

```

# How to put your data & metadata into one dataframe:  
Your metadata columns might have different names, based on the metadata you entered into CoralNet. Here, we give an example of metadata that includes site and transect, as these are commonly used metrics for analyzing data  

```{r main}


## Join together the data & metadata .csv outputs from CoralNet:
pointcount <- dplyr::left_join(data, metadata, by ="Name")


## Add functional group column to pointcount dataset, using labelset provided by CoralNet (posted on github)
pointcount <- dplyr::left_join(pointcount, labelset, by ="Label")


## Tidy up the column names on the pointcount dataframe:
pointcount<- janitor::clean_names(pointcount)


# Take a look at your dataframe:
head(pointcount)
```

# Summarizing your pointcount data:
```{r summarize}

# Quick way to get a frequency count on a column of interest:
plyr::count(pointcount$label)

#Count number of unique species in your dataset:

unique(pointcount$full_name)

```

## Calculate percent cover, using labels:  
* Across all sites  
* % cover by site  
* % cover by site and transect  

```{r labels}

# Percent cover of all labels, across all sites in your pointcount dataset:
perc_cov_lab<-pointcount %>%
  dplyr::group_by(label) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)

# Percent cover of each label, for each site in the dataset:
perc_cov_lab_site<- pointcount %>%
  dplyr::group_by(site, label) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)

# Percent cover of each label, by site & transect 
perc_cov_lab_site_trans<-pointcount %>%
  dplyr::group_by(site, transect, label) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)


```

## Calculate percent cover, based on functional groups:
* Across all sites  
* % cover by site  
* % cover by site and transect  
```{r functional}

# Percent cover of all functional groups, across all sites in your pointcount dataset:

perc_cov_funct<- pointcount %>%
  dplyr::group_by(functional_group) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)


# Percent cover of each functional group, for each site in the dataset:
perc_cov_funct_site<-pointcount %>%
  dplyr::group_by(site, functional_group) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)

# Percent cover of each functional group, by site & transect
perc_cov_funct_site_trans<- pointcount %>%
  dplyr::group_by(site, transect, functional_group) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(perc_cov = n / sum(n)*100)
```


